MCC Models

### 2019-04-15 papers/cogsci2019/cogsci2019.tex
kids model prediction with replication data
clean up repo

% KM added 1/4/18 to allow control of blind submission

@@

Participants selected the less frequent object above chance in the test
condition (t(39) = 5.51, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotexp12}) and did so more often compared to the control
condition (generalized linear mixed model
(GLMM\footnote{All models included random effects for subject and speaker.}):\emph{\(\beta\)}
= 1.49, se = 0.5 \emph{p} = .003). This result replicates earlier work
(Frank \& Goodman, 2014) and is consistent with the hypothesis that
listeners expect speakers to communicate in an informative way.

@@

Participants selected the preferred object above chance when the same
animal returned (t(39) = 29.14, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotexp12}) and did so more often compared to trials in which a
different animal returned (GLMM: \(\beta\) = 2.97, se = 0.7, \emph{p}
\textless{} .001). Thus, listeners inferred the referent of the
utterance by considering previous interactions with the speaker.
Interestingly, participants transferred preference to some extent from
one animal to the other and selected the preferred object above chance
when a different animal returned (t(39) = 2.7, \emph{p} = .01). In sum,
this study shows that adults make comparable inferences to children
(Saylor et al., 2009).

The novel object was selected above chance when the same animal returned
(t(39) = 6.77, \emph{p} \textless{} .001) but not when a different one
appeared (t(39) = 1.49, \emph{p} = .144, see Fig. \ref{fig:plotexp12}).
Furthermore, the two conditions differed in the expected direction
(\(\beta\) = 1.7, se = 0.39, \emph{p} \textless{} .001). Thus, like
children (Akhtar et al., 1996), adults used their prior information
about the speaker to resolve ambiguity in the utterance.

@@

Results are discussed in the form of the proportion with which listeners
chose the more informative object (i.e., the object that would be the
more informative referent when only considering speaker general
expectations). For a comparison to chance within each unique condition
see Fig. \ref{fig:plotexp3}. Combinations of alignment and speaker
identity differed in how they influenced participants' responses (GLMM
model term: \texttt{alignment*speaker}; \(\beta\) = -2.92, se = 0.57,
\emph{p} \textless{} .001). Fig. \ref{fig:plotmodelcomp}A shows the mean
response in each unique condition compared to the pragmatics model.
Model predictions and data were highly correlated (r = 0.96, \emph{p}
\textless{} .001). Model fit was much better in the model taking into
account both types of expectations compared to the uniform prior (BF =
2e+79) or prior only model (BF = 1.8e+34). The inferred noise level in
the pragmatics model was 0.27 (95\% HDI: 0.21 - 0.34).

@@

In our model, we treated common ground expectations as equivalent to
more basic manipulations of contextual salience (e.g.~in Frank \&
Goodman, 2012) and did not explicitly model the social-cognitive
processes that give rise to these expectations. The interaction around
the object prior to the test event simply increased the probability that
this particular speaker will refer to the object subsequently. The same
change could be brought about if one of the objects would be made
perceptually more salient, for example by making it flash. In future
work, it would be interesting to explore ways to model common ground
expectations explicitly as well as to contrast perceptual and
interactional salience.

This work integrates different perspectives on the study of pragmatic
inference. Previous work focused either on general or speaker specific
expectations. The methodological approach taken here illustrates how
computational and experimental approaches can be used in conjunction to
explicate theories of language use and learning.

@@

Masked for peer review.

@@

@@

@@

@@

>>>>

% KM added 1/4/18 to allow control of blind submission
\cogscifinalcopy

@@

Participants selected the less frequent object above chance in the test
condition (t(39) = 5.51, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotexp12}) and did so more often compared to the control
condition (generalized linear mixed model
(GLMM\footnote{All models had maximal random effects structure conditional on model convergence.}):\emph{\(\beta\)}
= 1.28, se = 0.29 \emph{p} \textless{} .001). This result replicates
earlier work (Frank \& Goodman, 2014) and is consistent with the
hypothesis that listeners expect speakers to communicate in an
informative way.

@@

Participants selected the preferred object above chance when the same
animal returned (t(39) = 29.14, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotexp12}) and did so more often compared to trials in which a
different animal returned (GLMM: \(\beta\) = 2.92, se = 0.56, \emph{p}
\textless{} .001). Thus, listeners inferred the referent of the
utterance by considering previous interactions with the speaker.
Interestingly, participants transferred preference to some extent from
one animal to the other and selected the preferred object above chance
when a different animal returned (t(39) = 2.7, \emph{p} = .01). In sum,
this study shows that adults make comparable inferences to children
(Saylor et al., 2009).

The novel object was selected above chance when the same animal returned
(t(39) = 6.77, \emph{p} \textless{} .001) but not when a different one
appeared (t(39) = 1.49, \emph{p} = .144, see Fig. \ref{fig:plotexp12}).
Furthermore, the two conditions differed in the expected direction
(\(\beta\) = 6.27, se = 1.96, \emph{p} = .001). Thus, like children
(Akhtar et al., 1996), adults used their prior information about the
speaker to resolve ambiguity in the utterance.

@@

Results are discussed in the form of the proportion with which listeners
chose the more informative object (i.e., the object that would be the
more informative referent when only considering speaker general
expectations). For a comparison to chance within each unique condition
see Fig. \ref{fig:plotexp3}. Combinations of alignment and speaker
identity differed in how they influenced participants' responses (GLMM
model term: \texttt{alignment*speaker}; \(\beta\) = -2.64, se = 0.48,
\emph{p} \textless{} .001). Fig. \ref{fig:plotmodelcomp}A shows the mean
response in each unique condition compared to the pragmatics model.
Model predictions and data were highly correlated (r = 0.96, \emph{p}
\textless{} .001). Model fit was much better in the model taking into
account both types of expectations compared to the uniform prior (BF =
2e+79) or prior only model (BF = 1.8e+34). The inferred noise level in
the pragmatics model was 0.27 (95\% HDI: 0.21 - 0.34).

@@

A range probabilistic frameworks been used to model word learning (e.g.
Fazly, Alishahi, \& Stevenson, 2010; Frank, Goodman, \& Tenenbaum, 2009;
Xu \& Tenenbaum, 2007). RSA models differ from other approaches in that
they treat word learning as the outcome of a social reasoning process.
In contrast to models for cross-situational word learning (Fazly et al.,
2010; Frank et al., 2009), RSA models show how learning might occur in a
one shot scenario based on pragmatic reasoning alone. While the ad hoc
informativeness inference characteristic for RSA would be predicted by
the model of Xu and Tenenbaum (2007), in their work it follows from the
``size principle'' of generalization (Tenenbaum \& Griffiths, 2001) and
not from social reasoning. In cotrast to RSA, this approach does not
offer a straightforward way to incorporate other types of social
information such as expectations following from common ground.

We treated common ground expectations as equivalent to more basic
manipulations of contextual salience (e.g.~in Frank \& Goodman, 2012)
and did not explicitly model the social-cognitive processes that give
rise to these expectations. The interaction around the object prior to
the test event simply increased the probability that this particular
speaker will refer to the object subsequently. The same change could be
brought about if one of the objects would be made perceptually more
salient, for example by making it flash. In future work, it would be
interesting to explore ways to model common ground expectations
explicitly as well as to contrast perceptual and interactional salience.

Our work integrates different perspectives on the study of pragmatic
inference. Previous work focused either on general or speaker specific
expectations. The methodological approach taken here illustrates how
computational and experimental approaches can be used in conjunction to
explicate theories of language use and learning.

\vspace{1em}
\fbox{\parbox[b][][c]{7.3cm}{\centering Corresponding data and code are available at\ \url{https://github.com/manuelbohn/mcc}}}
\vspace{1em}

@@

MB received funding from the European Union's Horizon 2020 research and
innovation programme under the Marie Sklodowska-Curie grant agreement No
749229.

@@

\hypertarget{ref-fazly2010probabilistic}{}
Fazly, A., Alishahi, A., \& Stevenson, S. (2010). A probabilistic
computational model of cross-situational word learning. \emph{Cognitive
Science}, \emph{34}(6), 1017--1063.

@@

\hypertarget{ref-frank2009using}{}
Frank, M. C., Goodman, N. D., \& Tenenbaum, J. B. (2009). Using
speakers' referential intentions to model early cross-situational word
learning. \emph{Psychological Science}, \emph{20}(5), 578--585.

@@

\hypertarget{ref-tenenbaum2001generalization}{}
Tenenbaum, J. B., \& Griffiths, T. L. (2001). Generalization,
similarity, and bayesian inference. \emph{Behavioral and Brain
Sciences}, \emph{24}(4), 629--640.

@@

\hypertarget{ref-xu2007word}{}
Xu, F., \& Tenenbaum, J. B. (2007). Word learning as bayesian inference.
\emph{Psychological Review}, \emph{114}(2), 245.

### 2019-04-15 papers/full_paper/manuscript.Rmd
clean up repo

# model (maximally converging)
lm_ex2pref <- glmer(correct ~ condition +
              (1|id), 
              data = adult_ex2_preference_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r ex3 GLMM, cache=T}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex3_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r ex4 GLMMs, cache=T}
# strong

@@

```{r child ex3 GLMM, cache = T}
child_ex3_lm_data <- child_ex3_data%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

>>>>

# model (maximally converging)
lm_ex2pref <- glmer(correct ~ condition +
              (1|id) + (1|agent), 
              data = adult_ex2_preference_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r ex3 GLMM, cache=F}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex3_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r ex4 GLMMs, cache=F}
# strong

@@

```{r child ex3 GLMM, cache = F}
child_ex3_lm_data <- child_ex3_data%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))
